---
title: "More cleaning, missing values, and subscale scores"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
sad_data <- read_csv("total_data.csv")
```
## More cleaning
For our analyses, we actually won't need the BRIEF or PANAS questionnaires, so we can omit them.
```{r omit brief}
sad_data <- sad_data %>% 
  select(-starts_with("BRIEF")) %>%
  select(-contains("PANAS"))
names(sad_data)
```

### Demographic variables

Cleaning up demographic variables

```{r clean up demographic variables}
# Let's inspect the age variable for suspect entries
sad_data %>% 
  distinct(age) %>%
  pull(.)
```

Looks like we have some missing values, some who erroneously input their ID numbers as their age, a 15 year old, and someone who entered zero. We'll omit the person who reported their age as 15 and fill the rest of entries with NAs

```{r clean age variable}
# omit 15 year old
sad_data <- sad_data %>% filter(!age == 15)

# there are 110 cases with questionable age variables, let's omit them. 
sad_data <- sad_data %>% mutate(age = na_if(age, 7087106163),
                    age = na_if(age, 6306244561),
                    age = na_if(age, 0))
  
# all the remaining age entries look good
sad_data %>% distinct(age)
```

Now, let's inspect the ethnicity variable

```{r ethnicity variable}
sad_data %>%
  distinct(ethnicity)
```

Appears to be in order, we have the two possible responses and some missing values.

```{r race}
sad_data %>%
  distinct(race)
```

Race looks fine too, just some missing values.

## Other missing values

Let's determine the percentage of each variable that is missing

```{r percentage of missing responses for each variable}
map_dbl(sad_data, ~ {
  (sum(is.na(.x))/sum(!is.na(.x))) * 100
  })
```
This is good news - it appears that no more than %5.70 of any variable is missing, so we had a pretty good response rate. Since we have such a large N, we can just omit all cases that have missing values.

```{r omit all cases with missing values}
sad_data <- na.omit(sad_data)
```
As you can see, the above function now indicates that 0% of all variables are missing
```{r test that there is no missing data}
map_dbl(sad_data, ~ {
  (sum(is.na(.x))/sum(!is.na(.x))) * 100
  })
```

```{r load data in workspace, include = FALSE}
write_csv(sad_data, "total_data.csv")
```

```{r code for finding percentage that each race and ethnicity comprises of the final sample}
# race
map_dbl(table(sad_data$race), ~{
  .x / nrow(sad_data) * 100
})
# ethnicity
map_dbl(table(sad_data$ethnicity), ~ {
  .x / nrow(sad_data) * 100
})
```

