---
title: "Path Analyses"
author: "Ian Kahrilas"
date: "5/28/2019"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
knitr::opts_chunk$set(echo = TRUE)
sad_data <- read_csv("total_data.csv")
```

First, we will load in the lavaan package. Lavaan is a structural equation modeling package for R. We will be using its more basic facilities to conduct path analyses.

```{r lavaan, message=FALSE}
library(tidyverse)
library(lavaan)
```

Now, we will construct our model. We have two predictors, positive affect (MASQAD14) and negative affect (MASQAD8) and an outcome of depression severity (PHQDEP). This relationship is mediated by the three temporal domains of savoring: reminiscing (SBI_REM), savoring the moment (SAV_MOM), and anticipating (SBI_ANT). We also include the covariates of sex, worry, and anxious arousal.

```{r path analysis}
sad_model <- '# covariates
             reminiscing ~ sex + pswq_total + masq_aa
             savoring_moment ~ sex + pswq_total + masq_aa
             anticipating ~ sex + pswq_total + masq_aa
             phq_total ~ sex + pswq_total + masq_aa
             masq_pa ~ sex + pswq_total + masq_aa
             masq_na ~ sex + pswq_total + masq_aa
           # direct effects
             phq_total ~ c1*masq_pa
             phq_total ~ ci*masq_na
           # mediators
             reminiscing ~ a1*masq_pa
             savoring_moment ~ a2*masq_pa
             anticipating ~ a3*masq_pa
             phq_total ~ b1*reminiscing
             phq_total ~ b2*savoring_moment
             phq_total ~ b3*anticipating
             reminiscing ~ ai*masq_na
             savoring_moment ~ aii*masq_na
             anticipating ~ aiii*masq_na
           # indirect effects (a*b)
             pa_rem := a1*b1
             pa_mom := a2*b2
             pa_ant := a3*b3
             na_rem := ai*b1
             na_mom := aii*b2
             na_ant := aiii*b3
           # total effects
             pa_total:= c1 + (a1*b1) + (a2*b2) + (a3*b3)
             na_total := ci + (ai*b1) + (aii*b2) + (aii*b3)
           # exogenous variable correlations
             masq_pa ~~ masq_na
           # residual correlations
             anticipating ~~ savoring_moment + reminiscing
             savoring_moment ~~ reminiscing
'
set.seed(123)
fit <- sem(sad_model, data = sad_data, se = "bootstrap", bootstrap = 10000, meanstructure = TRUE)
summary(fit, fit.measures = TRUE, ci = TRUE, standardized = TRUE, rsquare = TRUE)
parms <- parameterEstimates(fit, boot.ci.type = "bca.simple", standardized = TRUE)
sad_rsquare <- lavInspect(fit, "rsquare")
# save parameter dataframe to working directory so that it can be accessed in the table compiler script.
write_csv(parms, "parms.csv")
```

```{r multivariate regressions}
# the following package adds standardized regression coefficients to lm output
library(lm.beta) # install.packages("lm.beta")
# multiple regression to establish individual relations between PA, NA, and AA with depression
multi_mod <- lm(phq_total ~ masq_pa + masq_na + masq_aa, data = sad_data_int)
summary(multi_mod)
lm.beta(multi_mod)
```

