---
title: "Univariate EDA for PER Project"
author: "Ian Kahrilas"
date: "2/10/2020"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(pander)
library(ggpubr)
library(car)
library(emmeans)
library(welchADF)
library(moments)
library(modelbased)
library(patchwork)
library(r2glmm)
# load in data
erp_mast_wide <- read_csv(here("data", "paper_two", "data_from_PER_R_project", "created_data", "erp_mast_no_lp.csv"))
erp_avr_wide <- read_csv(here("data", "paper_two", "data_from_PER_R_project", "created_data", "erp_avr_no_lp.csv"))
per_questionnaires <- read_csv(here("data", "paper_two", "data_from_PER_R_project", "created_data", "per_measures.csv"))
```

```{r averaging and questionnaire variable selection, include = FALSE}
lpp_elec <- c("B28") # look at 450 - 800 ms windows
epn_elec_right <- "B26" # 250 - 375 ms
epn_elec_left <- "A29" # 250 - 375 ms
epn_elec_avg <- c("A29", "B27")
N170_right <- "B26" # 160 - 200 ms
N170_left <- "A29" # 160 - 200 ms
N170_avg <- c("A29", "B27")
# front_left <- c("EXG2", "A3", "A9", "A10") # 1000 - 2000 ms
# front_right <- c("EXG1", "B2", "B7", "B8")
# select variables from questionnaire data set for analyses
measures_int <- per_questionnaires %>% 
  select(pid:valence, age, sex, handedness, ethnicity, Race, political_orientation, coffee_typical_day:typical_sleep_hours,
         anticipating:masq_aa, tmms_repair:depression)

LPP <- erp_mast_wide %>%
  select(all_of(lpp_elec), pid:prop_trials) %>% 
  filter(between(ms, 450, 800)) %>%
  pivot_longer(., cols = all_of(lpp_elec), names_to = "electrode", values_to = "mv") %>% 
  group_by(pid, block) %>% 
  summarize(LPP = mean(mv, na.rm = TRUE),
            prop_trials = mean(prop_trials, na.rm = TRUE))

EPN_right <- erp_avr_wide %>%
  select(all_of(epn_elec_right), pid:prop_trials) %>% 
  filter(between(ms, 250, 375)) %>%
  pivot_longer(., cols = all_of(epn_elec_right), names_to = "electrode", values_to = "mv") %>% 
  group_by(pid, block) %>% 
  summarize(EPN_right = mean(mv, na.rm = TRUE))

EPN_left <- erp_avr_wide %>%
  select(all_of(epn_elec_left), pid:prop_trials) %>% 
  filter(between(ms, 250, 375)) %>%
  pivot_longer(., cols = all_of(epn_elec_left), names_to = "electrode", values_to = "mv") %>% 
  group_by(pid, block) %>% 
  summarize(EPN_left = mean(mv, na.rm = TRUE))

N170_right <- erp_avr_wide %>% 
  select(all_of(N170_right), pid:prop_trials) %>% 
  filter(between(ms, 160, 200)) %>% 
  pivot_longer(., cols = all_of(N170_right), names_to = "electrode", values_to = "mv") %>% 
  group_by(pid, block) %>% 
  summarize(N170_right = mean(mv, na.rm = TRUE))

N170_left <- erp_avr_wide %>% 
  select(all_of(N170_left), pid:prop_trials) %>% 
  filter(between(ms, 160, 200)) %>% 
  pivot_longer(., cols = all_of(N170_left), names_to = "electrode", values_to = "mv") %>% 
  group_by(pid, block) %>% 
  summarize(N170_left = mean(mv, na.rm = TRUE))

# front_right <- erp_mast_wide %>% 
#   select(all_of(front_right), pid:prop_trials) %>% 
#   filter(between(ms, 1000, 2000)) %>% 
#   pivot_longer(., cols = all_of(front_right), names_to = "electrode", values_to = "mv") %>% 
#   group_by(pid, block) %>% 
#   summarize(front_right = mean(mv, na.rm = TRUE))
  
# front_left <- erp_mast_wide %>% 
#   select(all_of(front_left), pid:prop_trials) %>% 
#   filter(between(ms, 1000, 2000)) %>% 
#   pivot_longer(., cols = all_of(front_left), names_to = "electrode", values_to = "mv") %>% 
#   group_by(pid, block) %>% 
#   summarize(front_left = mean(mv, na.rm = TRUE))
```

Join all the tibbles together
```{r join all tibbles together, include=FALSE}
#join all LPP tibbles plus right frontal
per_erp <- left_join(LPP, EPN_right, by = c("pid", "block")) %>%
  left_join(EPN_left, by = c("pid", "block")) %>%
  left_join(N170_right, by = c("pid", "block")) %>%
  left_join(N170_left, by = c("pid", "block")) %>% 
  # left_join(front_right, by = c("pid", "block")) %>% 
  # left_join(front_left, by = c("pid", "block")) %>% 
  left_join(measures_int, by = c("pid", "block"))
# change variable classes for parsimonious exploriation and analysis
## pid as character type
per_erp$pid <- as.character(per_erp$pid)
## block as factor
per_erp$block <- as.factor(per_erp$block)
per_erp <- per_erp %>% 
  select(pid, block, prop_trials, everything())
```

## Questionnaire Descriptive Statisitcs

```{r descriptive stats on variables of interest}
per_erp %>% 
  ungroup() %>% 
  summarize("PA Mean" = mean(pos_affectivity, na.rm = TRUE),
            "PA SD" = sd(pos_affectivity, na.rm = TRUE),
            "NA Mean" = mean(neg_affectivity, na.rm = TRUE),
            "NA SD" = sd(neg_affectivity, na.rm = TRUE),
            "StM Mean" = mean(savoring_moment, na.rm = TRUE),
            "StM SD" = sd(savoring_moment, na.rm = TRUE)) %>% 
  pander(split.tables = 200, caption = "PA = Positive Affectivity, NA = Negative Affectivity, StM = Savoring the Moment")
```

## EEG Descriptive Statistics by Block

```{r descriptive statistics for ERP components}
per_erp %>% 
  group_by(block) %>% 
  summarize(LPP_mean = mean(LPP, na.rm = TRUE),
            LPP_sd = sd(LPP, na.rm = TRUE),
            EPN_left_mean = mean(EPN_left, na.rm = TRUE),
            EPN_left_sd = sd(EPN_left, na.rm = TRUE),
            EPN_right_mean = mean(EPN_right, na.rm = TRUE),
            EPN_right_sd = sd(EPN_right, na.rm = TRUE),
            N170_right_mean = mean(N170_right, na.rm = TRUE),
            N170_right_sd = sd(N170_right, na.rm = TRUE),
            N170_left_mean = mean(N170_left, na.rm = TRUE),
            N170_left_sd = sd(N170_left, na.rm = TRUE))
```

## Behavioral Rating Descriptive Statistics by Block

```{r descriptive statistics for behavioral ratings}
per_erp %>% 
  group_by(block) %>% 
  summarize(arousal_mean = mean(arousal, na.rm = TRUE),
            arousal_sd = sd(arousal, na.rm = TRUE),
            valence_mean = mean(valence, na.rm = TRUE),
            valence_sd = sd(valence, na.rm = TRUE),
            difficulty_mean = mean(difficulty, na.rm = TRUE),
            difficulty_sd = sd(difficulty, na.rm = TRUE))
```

## Exploration of Questionnaire Data

```{r histograms and qqplots, message = FALSE, warning = FALSE}
meas_by_pid <- per_erp %>% 
  group_by(pid) %>% 
  summarize(anticipating = mean(anticipating, na.rm = TRUE),
            savoring_moment = mean(savoring_moment, na.rm = TRUE),
            reminiscing = mean(reminiscing, na.rm = TRUE),
            sbi_tot = mean(sbi_tot, na.rm = TRUE),
            pos_affectivity = mean(pos_affectivity, na.rm = TRUE),
            neg_affectivity = mean(neg_affectivity, na.rm = TRUE),
            pos_affect = mean(pos_affect, na.rm = TRUE),
            neg_affect = mean(neg_affect, na.rm = TRUE),
            masq_pa = mean(masq_pa, na.rm = TRUE),
            masq_na = mean(masq_na, na.rm = TRUE),
            masq_aa = mean(masq_aa, na.rm = TRUE),
            tmms_repair = mean(tmms_repair, na.rm = TRUE),
            tmms_attention = mean(tmms_attention, na.rm = TRUE),
            tmms_clarity = mean(tmms_clarity, na.rm = TRUE),
            erq_reappraisal = mean(erq_reappraisal, na.rm = TRUE),
            erq_suppression = mean(erq_suppression, na.rm = TRUE),
            depression = mean(depression, na.rm = TRUE))

measure_hist_qq_fun <- function(var, name) {
hist <- meas_by_pid %>% 
  gghistogram(., var)
qq <- meas_by_pid %>% 
  ggqqplot(., var)
hist|qq + plot_annotation(title = name)
}

name <- c("Anticipating", "Savoring the Moment", "Reminiscing", "SBI Total", "Positive Affectivity",
           "Negative Affectivity", "Positive Affect", "Negative Affect", "MASQ PA", "MASQ NA",
           "MASQ AA", "TMMS Repair", "TMMS Attention", "TMMS Clarity", "ERQ Reappraisal",
           "ERQ Suppression", "Depression")

map2(names(meas_by_pid)[-1], name, ~ measure_hist_qq_fun(.x, .y))
```

## Histograms and QQplots for each block with EEG Components

```{r histograms and qqplots for LPP, EPN, and N170 for each block, warning = FALSE, message = FALSE}
hist_qq_fun <- function(trial, component) {
(per_erp %>% 
  filter(block == trial) %>% 
  select(component) %>% 
  gghistogram(., x = component) |
per_erp %>% 
  filter(block == trial) %>% 
  select(component) %>% 
  ggqqplot(., component)) +
  plot_annotation(title = paste(trial, component))
}

blocks <- unique(as.character(per_erp$block))
components <- names(per_erp)[4:8]
map(components, ~map(blocks, ~hist_qq_fun(trial = .x, component = .y), .y = .x))
```
 
Omit extreme cases indicated in the histograms and QQ plots, and regenerate histograms and QQ plots.

```{r examine extreme cases, warning=FALSE,message=FALSE}
per_erp$LPP[per_erp$block == "Neg_Watch" & per_erp$LPP > 20] <- NA
per_erp$LPP[per_erp$block == "Pos_Watch" & per_erp$LPP < -5] <- NA

map(components, ~map(blocks, ~hist_qq_fun(trial = .x, component = .y), .y = .x))

# write new file to workspace
write_csv(per_erp, here("data", "paper_two", "created_data", "per_data_analyses_2020_5_19.csv"))
```

## Histograms and QQplots for each block with behavioral ratings

```{r hist and qq with behavioral ratings, echo = FALSE, message = FALSE, warning = FALSE}
hist_qq_fun <- function(trial, rating) {
(per_erp %>% 
  filter(block == trial) %>% 
  select(rating) %>% 
  gghistogram(., x = rating, bins = 8) |
per_erp %>% 
  filter(block == trial) %>% 
  select(rating) %>% 
  ggqqplot(., rating)) +
  plot_annotation(title = paste(c(trial, rating)))
}
blocks <- unique(as.character(per_erp$block))
ratings <- names(per_erp)[9:11]

map(ratings, ~map(blocks, ~hist_qq_fun(trial = .x, rating = .y), .y = .x))
```
